openarchaeoCollaboration

### 2021-06-14 analysis/paper.Rmd
Expand repo usage/host analysis
Also revise the language analysis text, to reflect that new platform analysis

```{r dependencies, include=FALSE}
library("tidyverse")
library("lubridate")
library("openarchaeoCollaboration")
library("khroma")
```

@@

#### Categories, tags, platforms and hosts

@@

The vast majority of projects included in open-archaeo are hosted on GitHub, but a few other hosts are used too. If we break this down by number of unique project maintainers, GitHub still remains dominant, but we see that each other platform is only used by a handful of individuals.

```{r plot-repo-hosts}
# TODO: add "none"
# TODO: version with unique authors
oarch %>% 
  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg),
               names_to = "repo_host", values_to = "repo_url") %>% 
  drop_na(repo_url) %>% 
  count(repo_host) %>% 
  arrange(-n) %>%
  ggplot(aes(fct_reorder(repo_host, n, .desc = TRUE), n)) +
  geom_col() +
  labs(title = "Repository hosts",
       caption = "Data: open-archaeo.info",
       x = NULL, y = "Number of projects")
```

ZB: Plot host by number of unique users.

How many are explicitly licensed? What are the most popular licenses?

@@

<<< cut? >>>

@@

TODO: colour by platform

@@

We also have a 'platform' taxonomy, which is a manually-classified description of the broad framework that the individual projects rely on.

<<< /cut? >>>

>>>>

```{r dependencies, include=FALSE}
library("tidyverse")
library("lubridate")
library("openarchaeoCollaboration")
library("khroma")
library("gt")
```

@@

#### Categories, tags, and platforms

@@

@@

@@

TODO: Do we need to use this data, given it says much the same as the 'platform' taxonomy?

@@

This largely confirms the pattern seen in our manual classification of the 'platform' of each project.

#### Version control

All projects included in open-archaeo make their source code available in some form, by definition.
But not all make use of version control.

```{r plot-repo-hosts}
oarch <- mutate(oarch, none = ifelse(is.na(github) & 
                                       is.na(gitlab) & 
                                       is.na(bitbucket) & 
                                       is.na(launchpad) & 
                                       is.na(codeberg),
                                     "", NA))

oarch %>% 
  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
               names_to = "repo_host", values_to = "repo_url") %>% 
  drop_na(repo_url) %>% 
  group_by(repo_host, author1_name) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  mutate(total = sum(n)) %>% 
  ungroup() %>% 
  mutate(repo_host = fct_reorder(repo_host, total),
         repo_host = fct_relevel(repo_host, "none")) %>% 
  ggplot(aes(repo_host, n, group = author1_name)) +
  geom_col(colour = "#333333") +
  coord_flip() +
  labs(title = "Most popular repository hosts",
       subtitle = "By number of projects, grouped by first author",
       caption = "Data: open-archaeo.info",
       x = NULL, y = NULL)
```

Amongst those that do use version control, the vast majority are hosted on GitHub, but a few other hosts are used too. Codeberg is a distant second-place in terms of popularity, but this is largely driven by a single active author.

```{r table-no-repo}
filter(oarch, none == "") %>% 
  count(category)
```

About one in eight projects do not use one of these repository host.
Some may use a different host, or not make their repositories publicly available, but it is probably safe to assume that the majority simply do not use version control.
Many of these are standalone documents or code fragments (scripts) hosted, for example, as 'gists', but there are also some pieces of standalone software that apparently do not include version control.
